---
title: "Hw3 for ANA 515"
author: "Pin-Chieh Huang"
date: "2024-06-20"
output:
 
  word_document: word_document
  html_document: html_document
---

Describe Data:

The database I am using is from United Airlines how much they spent on each department every year based on wide and narrow body aircraft. The data is stored as an Excel file. As we know most of data is about the money and number,so the data type is binary. In order to have perfect number, we need to  remove n/a row and convert the scientific number. In the report, we will see only summary and clean data from our file. 

```{r section 2~3, include = FALSE}
#install.packages("tidyverse")
#install.packages("readxl")
#install.packages("dplyr")
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(knitr)

library(magrittr)

united <- read_xls("United Airlines Aircraft Operating Statistics- Actuals.xls")
united
united1 <-united[-1,-1] # remove the first row and first column because of n/a value
colnames(united1) <- united1[1,] # to see the first row which is years
united2 <- united1[-1,] # remove the first now because we want to bring the first row (year) as head 
united2 <- na.omit(united2) #remove the rows with n/a
united3 <- subset(united1,select = c("year","2013","2014","2015"))
```



We have data about `r nrow(united3)` factories such as how much they spent on pilot, HR, aircrafts  and so on . The distribution of this is shown below: The data is about 2013,2014 and 2015 how much united airlines spent on the each fleet and pilot the `1column(united3)`is the data of 2013 and the `2column(united3)`is the data of 2014 and `3column(united3)`is the data of 2015


```{r section4, include=FALSE}
united2 <- subset(united1,select = c("year","2014","2015")) #subset 2014 and 2015 columns 
options(scipen = 999) #  to turn off scientific notation

united2 <- na.omit(united2) #remove the rows with n/a
print(kable(united2)) # the total costs in 2014 and 2015 from united differetn departments 
united2
```

```{r section5, eval=FALSE, include=FALSE}
united3 <- subset(united1,select = c("year","2013","2014","2015")) #subset 2013, 2014 and 2015 columns 
options(scipen = 999) #  to turn off scientific notation

united3 <- na.omit(united3) #remove the rows with n/a
print(kable(united3))
united3
united3_firstcolumn = united3[-1,+1]
united3_firstcolumn
colnames(united3_firstcolumn)<-c("category")
new_numited <- cbind(united3_firstcolumn, united3_1)
```


```{r echo=FALSE, comment=NA}

#install.packages("reshape2")
library(reshape2)
united3 <- subset(united1,select = c("year","2013","2014","2015")) #subset 2013, 2014 and 2015 columns 
options(scipen = 999) #  to turn off scientific notation

united3 <- na.omit(united3) #remove the rows with n/a

filterdate<-( filter(united3,year == "Small Narrowbodies" | year == "Total Fleet" ))
filterdate<-t(filterdate)


data <- data.frame(
  year = c(2013, 2014, 2015),
  total_fleet = c(17157931, 16996369, 13876654),
  Small_Narrowbodies = c(3296735, 3200531, 2409722))
data$year <- as.numeric(as.character(data$year))
# Plot the data
ggplot(data, aes(x = as.factor(year), y = total_fleet)) +
  geom_line(group = 1) +  # Added group = 1 for proper line plot
  geom_point() +
  geom_text(aes(label = total_fleet), vjust = -0.5) +
  labs(title = "Total Fleet Over the Years",
       x = "Year",
       y = "how much united spent on total Fleet") +
  scale_x_discrete(breaks = c(2013, 2014, 2015)) +  # Ensure only the years are shown
  theme_minimal()

# Reshape the data for the second plot
data_long <- melt(data, id = "year")

# Plot both total fleet and small narrowbodies over the years with points and data labels
ggplot(data_long, aes(x = year, y = value, color = variable)) +
  geom_line() +
  geom_point() +
  geom_text(aes(label = value), vjust = -0.5) +
  labs(title = "Fleet Data Over the Years",
       x = "Year",
       y = "how much united spent on samll aircraft",
       color = "Category") +
  scale_x_continuous(breaks = c(2013, 2014, 2015)) +  # Ensure only the years are shown
  theme_minimal()

```

```{r eval=FALSE, comment=NA, include=FALSE}
# Load necessary libraries
library(ggplot2)
library(reshape2)

# Create the data frame
data <- data.frame(
  year = c(2013, 2014, 2015),
  total_fleet = c(17157931, 16996369, 13876654),
  small_narrowbodies = c(3296735, 3200531, 2409722),
  widebody = c(7045375,	7158573, 5930869
)
)

data$total_fleet_pct_diff <- c(NA, diff(data$total_fleet) / head(data$total_fleet, -1) * 100)
data$small_narrowbodies_pct_diff <- c(NA, diff(data$small_narrowbodies) / head(data$small_narrowbodies, -1) * 100)

# Melt the data to long format for plotting
data_long <- melt(data, id = "year", measure.vars = c("total_fleet", "small_narrowbodies"))

# Melt the percentage difference data to long format for plotting
data_pct_diff_long <- melt(data, id = "year", measure.vars = c("total_fleet_pct_diff", "small_narrowbodies_pct_diff"))

# Plot the data
p1 <- ggplot(data_long, aes(x = year, y = value, color = variable)) +
  geom_line() +
  geom_point() +
  labs(title = "Fleet Data Over the Years",
       x = "Year",
       y = "Count",
       color = "Category") +
  theme_minimal()

# Plot the percentage difference data
p2 <- ggplot(data_pct_diff_long, aes(x = year, y = value, color = variable)) +
  geom_line() +
  geom_point() +
  labs(title = "Percentage Difference Over the Years",
       x = "Year",
       y = "Percentage Difference",
       color = "Category") +
  theme_minimal()

# Print the plots
p1
p2

ggplot(data = data, mapping = aes(x =year , colour = widebody)) +
  geom_freqpoly(binwidth = 0.1)

```




```{r pilot, echo=FALSE, message=FALSE}
library(data.table)
long <- melt(setDT(united3), id.vars = c("year"), variable.name = "year_new") #reshap dataframe
long

long$year_new <- as.numeric(as.character(long$year_new)) # Convert the year column to numeric



filtered_data_second <- long %>%
  filter( year_new >= 2013 & year== "Pilot Training (000)" | year=="Pilots (000)")

filtered_data_second

ggplot(filtered_data_second, aes(x = (year_new), y = value, fill = year)) +
  geom_bar(stat = "identity") +
  labs(title = "how much unied spent on pilot",
       x = "Year",
       y = "$$",
       fill = "$$") +
  theme_minimal()

```